A performance assessment of evolutionary algorithms in volunteer computing environments: the importance of entropy

### 2016-11-15 EvoPar2017/paper.Rnw
typo spotted
Couple of questions for @JJ related with the conclusion section. Related with issue #127

This is interesting in the sense that in indicates that there might be
something more to performance that just the number of volunteers
contributing. In this experiment we include an additional measure:
user-contribution entropy, which is measured by computing the
compression entropy of every user contribution for the whole %please check 
experiment.  This is computed converting the sequence of the number of
PUTs into a string and then compressing it using {\tt Zlib}, a
standard compression library. Since the compression rate is related to
the randomness of the sequence, a higher compression rate will
correspond to an experiment with a low entropy, while a lower
compression rate will indicate that there has been a certain amount of
unexpected changes during the experiment. Although less popular than
Jenssen-Shannon entropy, this way of measuring it is widely used
\cite{ornstein1993entropy} in case where we want to compute the
entropy of sequences of events, such as arrhythmias
\cite{baumert2004forecasting}. Unlike Jenssen-Shannon entropy, the
range of the compression entropy will vary between 0 (lower, or worse
compression rate) and 1 (no compression).

@@

Our main intention in this paper was to configure an evolutionary
algorithm in a volunteer computing environment to improve its
efficiency and eventually find out on what this efficiency depends
on. The baseline configuration that we had used in other papers 
\cite{jj-ppsn98anonymous} included
{\em overlaps} or volunteers running islands that had actually started in
a previous experiment instance in the set. In previous papers, we were
interested in measuring the number of users or their behavior in bulk
and this overlap was not so important, but
in this case our focus was on the evolutionary algorithm so we had to
make all volunteers in it started more or less at the same time and
from a random population. In fact, we managed to do so, except in some
byzantine cases, by making the clients restart when the cache size
collapses to 1 or to a size much smaller than the current; we called
this client configuration {\tt reboot} and {\tt cachecrash}.  This
configuration achieves a positive secondary effect: all clients
contribute to fill the cache, having thus an actual effect on the
chromosome pool and effectively making all clients present contribute
to the simulation.

The last configuration managed to find, with a relatively small number
of evaluations, solutions in half the time that in previous sets of
experiments. Although in the {\tt base} and {\tt graph} the median
time and number of contributions was smaller, we cannot disregard the
possibility that many of those solutions were found thanks to
contributions from {\em carried over} islands with an algorithm that
has run a big amount of evaluations already. At any rate, the {\tt
  cachereboot} finds on average solutions four times as fast as {\tt
  graph2} with twice as many average volunteers (or three times as
many if we consider {\em effective} volunteers), needing 50\% less
evaluations, thus meaning a big improvement in the performance of the
evolutionary algorithm, which was the target of this paper. This
improvement is less with respect to {\tt reboot}, but in any case the
method to detect the beginning of a new run is much more effective and
will be kept in the future. But the fact that this happened was due to
the {\em entropy} as proved in the last section of this paper.

>>>>

This is interesting in the sense that it indicates that there might be
something more to performance that just the number of volunteers
contributing. In this experiment we include an additional measure:
user-contribution entropy, which is measured by computing the
compression entropy of every user contribution for the whole %please check 
experiment.  This is computed converting the sequence of the number of
PUTs into a string and then compressing it using {\tt Zlib}, a
standard compression library. Since the compression rate is related to
the randomness of the sequence, a higher compression rate will
correspond to an experiment with a low entropy, while a lower
compression rate will indicate that there has been a certain amount of
unexpected changes during the experiment. Although less popular than
Jenssen-Shannon entropy, this way of measuring it is widely used
\cite{ornstein1993entropy} in case where we want to compute the
entropy of sequences of events, such as arrhythmias
\cite{baumert2004forecasting}. Unlike Jenssen-Shannon entropy, the
range of the compression entropy will vary between 0 (lower, or worse
compression rate) and 1 (no compression).

@@

Our main intention in this paper was to configure an evolutionary
algorithm in a volunteer computing environment to improve its
efficiency and eventually find out on what this efficiency depends
on. The baseline configuration that we had used in other papers 
\cite{jj-ppsn98anonymous} included
{\em overlaps} or volunteers running islands that had actually started in
a previous experiment instance in the set. In previous papers, we were
interested in measuring the number of users or their behavior in bulk
and this overlap was not so important.
In this case our focus was on the evolutionary algorithm so we had to
make all volunteers in it started more or less at the same time and
from a random population. %Pablo: too long paragraph, added commas and sepparated into two
In fact, we managed to do so, except in some
byzantine cases, by making the clients restart when the cache size
collapses to 1 or to a size much smaller than the current; we called
this client configuration {\tt reboot} and {\tt cachecrash}.  This
configuration achieves a positive secondary effect: all clients
contribute to fill the cache, having thus an actual effect on the
chromosome pool and effectively making all clients present contribute
to the simulation.

The last configuration managed to find, with a relatively small number
%Pablo: what is the difference with this configurarion with respect to the others?
of evaluations, solutions in half the time that in previous sets of
experiments. Although in the {\tt base} and {\tt graph} the median
%Pablo: I usually prefer to do self-contained and conclusions, this whole explanation could be moved to discussion as it is too specific. But it is up to you :)
time and number of contributions was smaller, we cannot disregard the
possibility that many of those solutions were found thanks to
contributions from {\em carried over} islands with an algorithm that
has run a big amount of evaluations already. At any rate, the {\tt
  cachereboot} finds on average solutions four times as fast as {\tt
  graph2} with twice as many average volunteers (or three times as
many if we consider {\em effective} volunteers), needing 50\% less
evaluations, thus meaning a big improvement in the performance of the
evolutionary algorithm, which was the target of this paper. This
improvement is less with respect to {\tt reboot}, but in any case the
method to detect the beginning of a new run is much more effective and
will be kept in the future. But the fact that this happened was due to
the {\em entropy} as proved in the last section of this paper.
